Policy optimization for dynamic spatiotemporal systems \hspace{1em} \insertframenumber

### 2014-11-21 pot/LaTeX/skewnormal.tex
tweaks to results

\section{Spatial skew processes}\label{s:skewt}
Many types of data demonstrate some level of skewness and therefore should be modeled with distributions that allow for non-symmetry.
The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter $\alpha$ to account for non-symmetric data \citep{Genton2004}.

\subsection{Skew-normal process}
Let $U(\bs)$ be a normalized skew-normal process with slant $\alpha$. We write $U(\bs) \sim \text{SN}_d(\Omega, \alpha)$, and the density is given as
\begin{align}
  2 \phi_d (U(\bs); \Omega) \Phi(\alpha \bOne^T U(\bs))
\end{align}
where $\phi_d$ is the density function for a $N_d(0, \Sigma)$ random variable, and $\Phi$ is a standard normal cumulative distribution function.
There is an additive representation \citep{Zhang2010} to construct a standardized skew-normal process given by
\begin{align}
  U(\bs) = \alpha |z| + v(\bs)
\end{align}
where $\alpha \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, and $v(\bs)$ is a Gaussian process with mean zero, variance one, \Matern spatial correlation with spatial range $\rho$, smoothness $\nu$, and $\gamma$ is the proportion of variance accounted for by the spatial variation.
So,
\begin{align}
  \text{cor}(U(\bs), U(\bs)) = \gamma I(s = t) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ ||\bs - \bt|| }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ ||\bs - \bt|| }{ \rho } \right).
\end{align}
Let $Y(\bs) = \bX(\bs)^T + \sigma U(\bs)$.
Then, after marginalizing over the $z$ term, we find that
\begin{align}
  Y(\bs) \sim \text{SN}_d(\bX^T(\bs)\bbeta, \Omega, \alpha)
\end{align}
where $\Omega = \sigma^2 (\Sigma + \alpha^2 \bOne \bOne^T$).

\subsection{Skew-$t$ process}
If $\sigma^2 \sim \text{IG}(a, b)$, after over the $\sigma^2$ terms, we have a skew-$t$ process with distribution function

\begin{align*}
  \bY \sim \text{skew-}t (\mu, \bSigma^*, \alpha, \text{df} = 2a)
\end{align*}
where $\mu$ is the location, $\bSigma^* = \frac{ b }{ a } \bSigma$, $\bSigma$ is a \Matern covariance matrix, and $\alpha \in \calR$ controls the skewness.
The skew-$t$ process is desirable because of its flexible tail that is controlled by the skewness parameter, $\alpha$, and the degrees of freedom, $2a$.
Furthermore, if $\bY$ follows a multivariate skew-$t$ distribution, the marginal distributions also follow a skew-$t$ distribution \citep{Azzalini2003}.

One member of this family, is the skew-$t$ distribution \citep[see Appendix A.5]{Azzalini2003}.
\citet{Zhang2010} show that the skew-$t$ process can be expressed as a hierarchical model
\begin{align} \label{eq:fullmodel}
  Y(\bs) &= \bX(\bs)^T \bbeta + \alpha |z| + \sigma v(\bs)
\end{align}
where $\alpha \in \calR$ controls the skewness; $z \sim N(0, \sigma^2)$ is a normal random effect; $v(\bs)$ is a Gaussian process with mean zero, variance one, \Matern spatial correlation with spatial range $\rho$, smoothness $\nu$, and $\gamma$ is the proportion of variance accounted for by the spatial variation; and $\sigma^2 \sim \text{IG}(a, b)$.
So,
\begin{align}
  \text{cor}(Y(\bs), Y(\bs)) = \gamma I(s = t) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ ||\bs - \bt|| }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ ||\bs - \bt|| }{ \rho } \right)
\end{align}

\subsection{Extremal dependence}
One measure of extremal dependence is the $\chi$ statistic \citep{Padoan2011}.
The $\chi$ statistic for the upper tail is given by
\begin{align*}
  \chi = \lim_{c \rightarrow \infty} \Pr(Y(\bs_1) > c | Y(\bs_2) > c).
\end{align*}
In a stationary spatial process, we can write the $\chi$ coefficient as
\begin{align*}
  \chi(h) = \lim_{c \rightarrow \infty} \Pr(Y(0) > c | Y(h) > c)
\end{align*}
where $h = ||\bs_1 - \bs_2||$.
If $\chi(h) = 0$, then observations are asymptotically independent at distance $h$.
For Gaussian processes, $\chi(h) = 0$ regardless of the the distance, so they are not suitable for modeling spatially dependent extremes.
However, for the skew-$t$ process described in \ref{s:skewt}, $\chi(h) > 0$ \citep[they give an expression that is the sum of two survival functions from the skew-$t$ distribution. The expression isn't terribly informative, but I can include it if you think it would be helpful.]{Padoan2011}.

\section{Spatiotemporal skew-$t$ model for extremes}\label{s:spatial}
In this section, we describe the

@@

\subsection{Partitioning to remove long-range asymptotic dependence}\label{s:part}
One problem with the spatial skew-$t$ process is that all sites are asymptotically dependent regardless of their spatial separation.
This occurs because all observations, both near and far, share the same $z$ and $\sigma^2$ terms.
Even more problematic is that $\lim_{h \rightarrow \infty} \chi(h) \neq 0$.
We handle this problem with a random partition similar to \citet{Kim2005} that allows $z$ and $\sigma^2$ to vary by site.

Consider a set of spatial knots $\bw_{k} \sim \text{Uniform}(\calD)$ that define a random daily partition $P_{1}, \ldots, P_{K}$ of the spatial domain of interest $\calD \subset \calR^2$ such that
\begin{align*}
  P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \}.
\end{align*}
Then, the observation at each site becomes
\begin{align}
  Y(\bs) &= X(\bs) \bbeta + \alpha |z(\bs)| + \sigma v(\bs),
\end{align}
and, for $\bs \in P_{k}$,
\begin{align}
  z(\bs) &= z_{tk} \label{eq:sitez}\\
  \sigma(\bs) &= \sigma_{k} \label{eq:sitesig}.
\end{align}
So, within each partition, $\bY$ follows the distribution given in (\ref{eq:fullmodel}).

When incorporating the random daily partition, for two sites in the same partition, the $\chi$ statistic is $\chi(h) = \chi_{\text{skew-}t}(h)$. However, for two sites in different partitions, the $\chi$ statistic should be the same as the Gaussian $\chi(h)$, because the only correlation between the two sites, is the correlation given by $\bSigma$.
So, to show that $\lim_{h \rightarrow \infty} \chi(h) = 0$, we need only know that $\lim_{h \rightarrow \infty} \pi(h) = 0$
where $\pi(h)$ is the probability that two sites separated by distance $h$ are in the same partition.
A proof of this is given in Appendix A.3.

@@

Then
\begin{align}
   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{tk}, \alpha, \beta, \Theta &= X_t(\bs) \beta + \alpha |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
   z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\\
   \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\\
   \alpha &\sim N(0, 10)\\
   v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
   z_{tk} \mid z_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2))\\
   \sigma^{2*}_{tk} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\\
   \bw^*_{tk} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2))
\end{align}
where $\Theta = \{\rho, \nu, \gamma\}$; \mbox{$k = \argmin_\ell ||\bs - \bw_\ell ||$}; and $\Sigma$ is a \Matern covariance matrix with variance one, spatial range $\rho$, smoothness $\nu$, and the proportion of variance accounted for by the spatial variation is $\gamma$.

@@

We can use Gibbs sampling to update $Y_t(\bs)$ for censored observations that are below the threshold $T$.
After conditioning on $\alpha$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals.
So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\mu(\bs), \bSigma)$.
After imputing the censored observations, we update the model parameters.
To update the model parameters, we use standard Gibbs updates for parameters when possible.
In the case Gibbs sampling is not possible, parameters are updated using a random-walk Metropolis Hastings algorithm.
See Appendices A.1 and A.2 for details regarding the MCMC.
The final step of the computation is to use Bayesian Kriging to generate a predictive distribution for $Y_t(\bs^*)$ at prediction location $\bs^*$.
This step is similar to the imputation for censored observations except that the full conditionals are no longer truncated at $T$.

@@

\subsection{Design}\label{s:simdesign}
For all simulation designs, we generate data from the model in Section \ref{s:part} using $n_s=144$ sites and $n_t=50$ independent days.
The sites are generated Uniform$([0, 10] \times [0, 10])$.
We generate data from 6 different simulation designs:
\begin{enumerate} \setlength{\itemsep}{-0.5em}
  \item Gaussian marginal, $K=1$ knot
  \item $T$ marginal, $K=1$ knot
  \item $T$ marginal, $K=5$ knots
  \item Skew-$t$ marginal, $K=1$ knots
  \item Skew-$t$ marginal, $K=5$ knots
  \item Max-stable.
\end{enumerate}
In the first five designs, the $v_t(\bs)$ terms are generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\rho = 0.1$.
For the covariance matrices in designs 1 -- 5, the proportion of the variance accounted for by the spatial variation is $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect is $0.1$.
In the first design, $\sigma^2 = 2$ is used for all days.
For designs 2 -- 4, $\sigma^2_{tk} \iid \text{IG}(3, 8)$
For designs 1 -- 3, we set $\alpha = 0$.
For designs four and five, $\alpha = 3$ was used, and the $z_t$ are generated as described in (\ref{eq:sitez}).
In the sixth design, we generate from a spatial max-stable distribution \citep{Reich2012} with parameters $\mu = 1, \sigma=1, \xi=0.2$ and 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
\hl{I know I need more here about the max-stable method}
In all six designs, the mean $\mu(\bs) = 10$ is assumed to be constant across space.

$M = 50$ data sets are generated for each design.
For each data set we fit the data using
\begin{enumerate} \setlength{\itemsep}{-0.5em}
  \item Gaussian marginal, $K=1$ knots
  \item Skew-$t$ marginal, $K=1$ knots, $T=-\infty$
  \item $T$ marginal, $K=1$ knots, $T=q(0.80)$
  \item Skew-$t$ marginal, $K=5$ knots, $T=-\infty$
  \item $T$ marginal, $K=5$ knots, $T=q(0.80)$
\end{enumerate}
where $q(0.80)$ is the 80th sample quantile of the data.
The design matrix $\bX$ includes an the intercept with a prior of $\beta \sim \text{N}(0, 10)$.
The spatial covariance parameters have priors $\log(\nu) \sim \text{N}(-1.2, 1)$, $\gamma \sim \text{U}(0, 1)$, $\log(\rho) \sim \text{N}(-2, 1)$.
The skewness parameter has prior $\alpha \sim \text{N}(0, 2)$.
The residual variance terms have priors $\sigma^2_t(\bs) \sim \text{IG}(0.1, 0.1)$.
The knots have priors $\bw \sim \text{Unif}(\calD)$.
We do not fit the data using the max-stable methods from \citet{Reich2012} because of the time it takes.

@@

\subsection*{Variance terms}
When there is more than one site in a partition, then we update $\sigma^2_{tk}$ using a Metropolis Hastings update.
First, we generate from one of two candidate distributions.
In the case of temporal dependence, we generate a candidate for ${\sigma^*_{tk}}^2$ using a random walk Gaussian candidate distribution
\begin{align*}
  {\sigma^{*2}_{tk}}^{(c)} \sim \text{N}({\sigma^{*2}_{tk}}^{(r - 1)}, s^2)
\end{align*}
where ${{\sigma^*_{tk}}^2}^{(r - 1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
In the case where there is no temporal dependence, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution in an independence Metropolis Hastings update where $a^* = (n_{tk} + 1) / 2 + a$, $b^* = [Y_{tk}^T \Sigma^{-1}_{tk} Y_{tk} + z_{tk}^2] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
The acceptance ratio is
\begin{align*}
  R = \left\{
    \frac{ l[Y_t(\bs) | {\sigma^2_{tk}}^{(c)}, \ldots] }
         { l[Y_t(\bs) | {\sigma^2_{tk}}^{(r - 1)}]}
    \right\} \times \left\{
    \frac{ l[z_tk | {\sigma^2_{tk}}^{(c)}, \ldots] }
         { l[z_tk | {\sigma^2_{tk}}^{(r - 1)}, \ldots] }
    \right\} \times \left\{
    \frac{ p[ {\sigma^2_{tk}}^{(c)} ] }
         { p[ {\sigma^2_{tk}}^{(r - 1)}] }
    \right\} \times \left\{
    \frac{ c[ {\sigma^2_{tk}}^{(r - 1)}] }
         { c[ {\sigma^2_{tk}}^{(c)}]}
    \right\}
\end{align*}
where $p[\cdot]$ is the prior either taken from the time series given in (\ref{s:temporal}) or assumed to be IG$(a, b)$, and $c[\cdot]$ is the candidate distribution.
The candidate is accepted with probability $\min\{R, 1\}$.

@@

\section*{Appendix A.2: Posterior distributions}
\input{zpost.tex}
\input{betapost.tex}
\input{sigpost.tex}
\input{alphapost.tex}

@@

\section*{Appendix A.5: Multivariate skew-$t$ distribution}
\input{mvskewt.tex}

\bibliographystyle{rss}
\bibliography{skewnormal}

>>>>

\section{Spatial skew processes}\label{s:spatialskew}
Many types of data demonstrate some level of skewness and therefore should be modeled with distributions that allow for non-symmetry.
The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter $\alpha$ to account for non-symmetric data \citep{Genton2004}.
A brief review of the additive process by which a skew-$t$ process is created is given here.

\subsection{Skew-normal process} \label{s:skewnorm}
A skew-normal process can be constructed using an additive representation \citep{Zhang2010} given by
\begin{align}
  U(\bs) = \lambda |z| + v(\bs)
\end{align}
where $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, and $v(\bs)$ is a Gaussian process with mean zero, variance one, \Matern spatial correlation with spatial range $\rho$, smoothness $\nu$, and $\gamma$ is the proportion of variance accounted for by the spatial variation.
So,
\begin{align}
  \text{cor}(v(\bs), v(\bt)) = \gamma I(s = t) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ ||\bs - \bt|| }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ ||\bs - \bt|| }{ \rho } \right).
\end{align}
Then, after marginalizing over the $z$ term, we find that
\begin{align}
  U(\bs) \sim \text{SN}_d(0, \bar{\bOmega}, \balpha)
\end{align}
where $\bar{\bOmega} = (\Sigma + \lambda^2 \bOne \bOne^T$), $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters, and SN is a skew-normal process with correlation matrix $\bar{\bOmega}$ \citep{Azzalini2013}.

\subsection{Skew-$t$ process} \label{s:skewt}
Let
\begin{align}
  Y(\bs) = \bX(\bs)^T \bbeta + \sigma \lambda |z| + \sigma \bv(\bs)
\end{align}
where $\sigma^2 \sim \text{IG}(a, b)$ is an inverse gamma random variable.
Then after marginalizing over both the $z$ and $\sigma$ terms, we find that
\begin{align}
  Y(\bs) \sim \text{ST}(\bX(\bs)^T \bbeta, \bOmega, \balpha, 2a) \label{eq:fullmodel}
\end{align}
where $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega}$ and $\balpha$ are given in Section \ref{s:skewnorm}, and ST$(\mu(\bs), \bOmega, \balpha, 2a)$ represents a $d$-dimensional skew-$t$ process with location $\mu(\bs)$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$.

This kind of process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
Furthermore, the marginal distributions also follow a skew-$t$ distribution \citep{Azzalini2013}.
If we partition a $d$-dimensional vector of observations $Y(\bs)$ into two components $Y(\bs_1)$ and $Y(\bs_2)$ of size $p$ and $d - p$ respectively, then after marginalizing over $Y(\bs_2)$, we get that
\begin{align}
  Y(\bs_1) \sim \text{ST}_p(\bX(\bs_1)^T \bbeta, \bOmega_{11}, \balpha_{1(2)}, 2a)
\end{align}
where
\begin{align}
  \bX(\bs) &= \left( \begin{array}{c}
    \bX(\bs_1) \\
    \bX(\bs_2)
  \end{array} \right), \quad
  \bOmega = \left( \begin{array}{cc}
    \bOmega_{11} & \bOmega_{12} \\
    \bOmega_{21} & \bOmega_{22}
  \end{array} \right), \quad
  \balpha = \left(\begin{array}{c}
    \balpha_1\\
    \balpha_2
  \end{array} \right),\\
  \balpha_{1(2)} &= (1 = \balpha_2^T \bar{\bOmega}_{22 \cdot 1} \balpha)^{-1/2}(\balpha_1 + \bar{\bOmega}_{11}^{-1} \bar{\bOmega}_{12} \balpha_{2}),\\
  \bar{\bOmega}_{11}^{-1} &= (\bar{\bOmega})^{-1}, \qquad \bar{\bOmega}_{22 \cdot 1} = \bar{\bOmega}_{22} - \bar{\bOmega}_{21} \bar{\bOmega}_{11}^{-1} \bar{\bOmega}_{12}.
\end{align}

\subsection{Extremal dependence}
One measure of extremal dependence is the $\chi$ statistic \citep{Padoan2011}.
The $\chi$ statistic for the upper tail is given by
\begin{align}
  \chi(\bs_1, \bs_2) = \lim_{c \rightarrow \infty} \Pr(Y(\bs_1) > c | Y(\bs_2) > c).
\end{align}
In a stationary spatial process, we can write the $\chi$ coefficient as
\begin{align}
  \chi(h) = \lim_{c \rightarrow \infty} \Pr(Y(0) > c | Y(h) > c)
\end{align}
where $h = ||\bs_1 - \bs_2||$.
If $\chi(h) = 0$, then observations are asymptotically independent at distance $h$.
For Gaussian processes, $\chi(h) = 0$ regardless of the the distance, so they are not suitable for modeling spatially dependent extremes.
One problem with the spatial skew-$t$ process is that all sites are asymptotically dependent regardless of their spatial separation.
This occurs because all observations, both near and far, share the same $z$ and $\sigma^2$ terms.
Even more problematic is that $\lim_{h \rightarrow \infty} \chi(h) > 0$ \citep[see Appendix A.5]{Padoan2011}.

\section{Spatiotemporal skew-$t$ model for extremes}\label{s:spatial}
In this section, we introduce censoring to focus on tail behavior, and a random partition to remove long-range asymptotic dependence.

@@

\subsection{Partitioning to remove long-range asymptotic dependence}\label{s:part}
We handle the problem or long-range asymptotic dependence with a random partition similar to \citet{Kim2005} for Gaussian data that allows $z$ and $\sigma^2$ to vary by site.
Then, the observation at each site becomes
\begin{align}
  Y(\bs) &= \bX(\bs)^T \bbeta + \lambda |z(\bs)| + \sigma v(\bs).
\end{align}
To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process with rate $\lambda$ over spatial domain $\calD \in \calR^2$
The knots define a random daily partition $P_{1}, \ldots, P_{K}$ of $\calD$ given by
\begin{align*}
  P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \},
\end{align*}
and, for $\bs \in P_{k}$,
\begin{align}
  z(\bs) &= z_{k} \label{eq:sitez}\\
  \sigma(\bs) &= \sigma_{k} \label{eq:sitesig}.
\end{align}
So, within each partition, $Y(\bs)$ follows the distribution given in (\ref{eq:fullmodel}).
Across partitions, the $Y(\bs)$ remain correlated via the correlation function for $v(\bs)$ because it spans the partitions.

When incorporating the random daily partition, the $\chi$ statistic is for two sites $\bs$ in partition $k_s$ and $\bt$ in partition $k_t$ is
\begin{align}
  \chi(h) = I(k_s = k_t) \chi_{\text{skew-}t}(h) + I(k_s \neq k_t) \chi_{\text{Gaus}}(h).
\end{align}
However, for two sites in different partitions, the $\chi$ statistic should be the same as the Gaussian $\chi(h)$, because the only correlation between the two sites, is the correlation given by $\bSigma$.
So, to show that $\lim_{h \rightarrow \infty} \chi(h) = 0$, we need only know that $\lim_{h \rightarrow \infty} \pi(h) = 0$
where $\pi(h)$ is the probability that two sites separated by distance $h$ are in the same partition.
A proof of this is given in Appendix A.3.

@@

Then
\begin{align}
   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{tk}, \alpha, \beta, \Theta &= \bX_t(\bs)^T \beta + \lambda |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
   z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\\
   \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\\
   \lambda &\sim N(0, 10)\\
   v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
   z_{tk} \mid z_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2))\\
   \sigma^{2*}_{tk} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\\
   \bw^*_{tk} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2))
\end{align}
where $\Theta = \{\rho, \nu, \gamma\}$; \mbox{$k = \argmin_\ell ||\bs - \bw_\ell ||$}; and $\Sigma$ is a \Matern covariance matrix with variance one, spatial range $\rho$, smoothness $\nu$, and the proportion of variance accounted for by the spatial variation is $\gamma$.

@@

We can use Gibbs sampling to update $Y_t(\bs)$ for censored observations that are below the threshold $T$.
After conditioning on $\lambda$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals.
So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\mu(\bs), \bSigma)$.
After imputing the censored observations, we update the model parameters.
To update the model parameters, we use standard Gibbs updates for parameters when possible.
In the case Gibbs sampling is not possible, parameters are updated using a random-walk Metropolis Hastings algorithm.
See Appendices A.1 and A.2 for details regarding the MCMC.
The final step of the computation is to use Bayesian Kriging to generate a predictive distribution for $Y_t(\bs^*)$ at prediction location $\bs^*$.
This step is similar to the imputation for censored observations except that the full conditionals are no longer truncated at $T$.

@@

\subsection{Design}\label{s:simdesign}
For all simulation designs, we generate data from the model in Section \ref{s:part} using $n_s=144$ sites and $n_t=50$ independent days.
The sites are generated Uniform$([0, 10] \times [0, 10])$.
We generate data from 6 different simulation designs:
\begin{enumerate} \setlength{\itemsep}{-0.5em}
  \item Gaussian marginal, $K=1$ knot
  \item $T$ marginal, $K=1$ knot
  \item $T$ marginal, $K=5$ knots
  \item Skew-$t$ marginal, $K=1$ knots
  \item Skew-$t$ marginal, $K=5$ knots
  \item Max-stable.
\end{enumerate}
In the first five designs, the $v_t(\bs)$ terms are generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\rho = 0.1$.
For the covariance matrices in designs 1 -- 5, the proportion of the variance accounted for by the spatial variation is $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect is $0.1$.
In the first design, $\sigma^2 = 2$ is used for all days.
For designs 2 -- 4, $\sigma^2_{tk} \iid \text{IG}(3, 8)$
For designs 1 -- 3, we set $\lambda = 0$.
For designs four and five, $\lambda = 3$ was used, and the $z_t$ are generated as described in (\ref{eq:sitez}).
In the sixth design, we generate from a spatial max-stable distribution \citep{Reich2012} with parameters $\mu = 1, \sigma=1, \xi=0.2$ and 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
\hl{I know I need more here about the max-stable method}
In all six designs, the mean $\mu(\bs) = 10$ is assumed to be constant across space.

$M = 50$ data sets are generated for each design.
For each data set we fit the data using
\begin{enumerate} \setlength{\itemsep}{-0.5em}
  \item Gaussian marginal, $K=1$ knots
  \item Skew-$t$ marginal, $K=1$ knots, $T=-\infty$
  \item $T$ marginal, $K=1$ knots, $T=q(0.80)$
  \item Skew-$t$ marginal, $K=5$ knots, $T=-\infty$
  \item $T$ marginal, $K=5$ knots, $T=q(0.80)$
\end{enumerate}
where $q(0.80)$ is the 80th sample quantile of the data.
The design matrix $\bX$ includes an the intercept with a prior of $\beta \sim \text{N}(0, 10)$.
The spatial covariance parameters have priors $\log(\nu) \sim \text{N}(-1.2, 1)$, $\gamma \sim \text{U}(0, 1)$, $\log(\rho) \sim \text{N}(-2, 1)$.
The skewness parameter has prior $\lambda \sim \text{N}(0, 2)$.
The residual variance terms have priors $\sigma^2_t(\bs) \sim \text{IG}(0.1, 0.1)$.
The knots have priors $\bw \sim \text{Unif}(\calD)$.
We do not fit the data using the max-stable methods from \citet{Reich2012} because of the time it takes.

@@

\subsection*{Variance terms}
When there is more than one site in a partition, then we update $\sigma^2_{tk}$ using a Metropolis Hastings update.
First, we generate from one of two candidate distributions.
In the case of temporal dependence, we generate a candidate for ${\sigma^*_{tk}}^2$ using a random walk Gaussian candidate distribution
\begin{align*}
  {\sigma^{*2}_{tk}}^{(c)} \sim \text{N}({\sigma^{*2}_{tk}}^{(r - 1)}, s^2)
\end{align*}
where ${{\sigma^*_{tk}}^2}^{(r - 1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
In the case where there is no temporal dependence, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution in an independence Metropolis Hastings update where $a^* = (n_{tk} + 1) / 2 + a$, $b^* = [Y_{tk}^T \Sigma^{-1}_{tk} Y_{tk} + z_{tk}^2] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
The acceptance ratio is
\begin{align*}
  R = \left\{
    \frac{ l[Y_t(\bs) | {\sigma^2_{tk}}^{(c)}, \ldots] }
         { l[Y_t(\bs) | {\sigma^2_{tk}}^{(r - 1)}]}
    \right\} \times \left\{
    \frac{ l[z_{tk} | {\sigma^2_{tk}}^{(c)}, \ldots] }
         { l[z_{tk} | {\sigma^2_{tk}}^{(r - 1)}, \ldots] }
    \right\} \times \left\{
    \frac{ p[ {\sigma^2_{tk}}^{(c)} ] }
         { p[ {\sigma^2_{tk}}^{(r - 1)}] }
    \right\} \times \left\{
    \frac{ c[ {\sigma^2_{tk}}^{(r - 1)}] }
         { c[ {\sigma^2_{tk}}^{(c)}]}
    \right\}
\end{align*}
where $p[\cdot]$ is the prior either taken from the time series given in (\ref{s:temporal}) or assumed to be IG$(a, b)$, and $c[\cdot]$ is the candidate distribution.
The candidate is accepted with probability $\min\{R, 1\}$.

@@

\section*{Appendix A.2: Posterior distributions}
\input{zpost.tex}
\input{betapost.tex}
\input{sigpost.tex}
\input{lambdapost.tex}

@@

\section*{Appendix A.5: Skew-$t$ distribution}
\subsection*{Univariate extended skew-$t$ distribution}
We say that $Y$ follow a univariate extended skew-$t$ distribution with location $\xi \in \calR$, scale $\omega > 0$, skew parameter $\alpha \in \calR$, extended parameter $\tau \in \calR$, and degrees of freedom $\nu$ if has distribution function
\begin{align}
  f_{\text{EST}}(y) = \omega^{-1} \frac{ f_T (z; \nu) }{ F_T(\tau / \sqrt{ 1 + \alpha^2}; \nu)} F_T\left( (\alpha z + \tau) \sqrt{ \frac{ \nu + 1 }{ \nu + z^2}}; 0, 1, \nu + 1 \right)
\end{align}
where $f_T(t; \nu)$ is a univariate Student's $t$ with $\nu$ degrees of freedom, $F_T(t; \nu) = P(T < t)$, and $z = (y - \xi) / \omega$.
In the case that $\tau = 0$, then $Y$ follows a univariate skew-$t$ distribution.
\subsection*{Multivariate skew-$t$ distribution}
If $\bZ \sim \text{ST}_d(0, \bar{\bOmega}, \balpha, \nu)$ is a $d$-dimensional skew-$t$ distribution, and $\bY = \xi + \bomega \bZ$, where $\bomega = \text{diag}(\omega_1, \ldots, \omega_d)$, then the density of $Y$ at $y$ is
\begin{align*}
  f_y(\by) = det(\bomega)^{-1} f_z(\bz)
\end{align*}
where
\begin{align*}
  f_z(\bz) &= 2 t_d(\bz; \bar{\bOmega}, \nu) T \left( \balpha^T \bz \sqrt{ \frac{\nu + d}{\nu + Q(\bz)} }; \nu + d\right) \\
  \bz &= \bomega^{-1}(\by - \xi)
\end{align*}
where $t_d(\bz; \bar{\bOmega}, \nu)$ is a $d$-dimensional Student's $t$-distribution with scale matrix $\bar{\bOmega}$ and degrees of freedom $\nu$, $Q(z) = \bz^T \bar{\Omega}^{-1}\bz$ and $T(\cdot; \nu)$ denotes the univariate Student's $t$ distribution function with $\nu$ degrees of freedom \citep{Azzalini2013}.

\subsection*{Extremal dependence}
The $\chi$ statistic \citep{Padoan2011} is given by
\begin{align}
  \chi = \bar{F}_{\text{EST}}\left\{ \frac{(x_1^{1 / \nu} - \omega) \sqrt{\nu + 1} }{\sqrt{1 - \omega^2}}; 0, 1, \alpha_1, \tau_1, \nu+1 \right\} + \bar{F}_{\text{EST}}\left\{ \frac{ (x_2^{1 / \nu} - \omega) \sqrt{\nu + 1} }{ \sqrt{1 - \omega^2} }; 0, 1, \alpha_2, \tau_2, \nu + 1 \right\},
\end{align}
where $\bar{F}_{\text{EST}}$ is the univariate survival extended skew-$t$ function with zero location, unit scale, $\alpha_j = \alpha_i \sqrt{1 - \omega^2}$, $\tau_j = \sqrt{\nu + 1}(\alpha_j + \alpha_i \omega)$, and $x_j = F_T(\bar{\alpha}_i \sqrt{\nu + 1}; 0, 1, \nu) / F_T(\bar{\alpha}_j \sqrt{\nu + 1}; 0, 1, \nu)$ with $j = 1, 2$ and $i = 2, 1$ and where $\bar{\alpha}_j = (\alpha_j + \alpha_i \omega) / \sqrt{ 1 + \alpha_i^2 (1 - \omega^2)}$
\bibliographystyle{rss}
\bibliography{skewnormal}
